---
title: "FARS accidents data"
author: "Adam Kosewski"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
This package reads csv files, processes and summarize data. 
Data comes from the US National Highway Traffic Safety Administration's Fatality Analysis Reporting System (FARS), which is a nationwide census providing the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes. 

```{r results='hide', message=FALSE, warning=FALSE, echo=FALSE}
library("dplyr")
library("tidyr")
library("maps")
library("graphics")

fars_read <- function(filename) {
  if(!file.exists(filename))
    stop("file '", filename, "' does not exist")
  data <- suppressMessages({
    readr::read_csv(filename, progress = FALSE)
  })
  dplyr::tbl_df(data)
}
make_filename <- function(year) {
  year <- as.integer(year)
  sprintf(file.path("..","data-raw","accident_%d.csv.bz2"), year)
}

fars_read_years <- function(years) {
  lapply(years, function(year) {
    file <- make_filename(year)
    tryCatch({
      dat <- fars_read(file)
      dplyr::mutate(dat, year = year) %>%
        dplyr::select(MONTH, year)
    }, error = function(e) {
      warning("invalid year: ", year)
      return(NULL)
    })
  })
}

fars_summarize_years <- function(years) {
  dat_list <- fars_read_years(years)
  dplyr::bind_rows(dat_list) %>%
    dplyr::group_by(year, MONTH) %>%
    dplyr::summarize(n = n()) %>%
    tidyr::spread(year, n)
}
fars_map_state <- function(state.num, year) {
  filename <- make_filename(year)
  data <- fars_read(filename)
  state.num <- as.integer(state.num)

  if(!(state.num %in% unique(data$STATE)))
    stop("invalid STATE number: ", state.num)
  data.sub <- dplyr::filter(data, STATE == state.num)
  if(nrow(data.sub) == 0L) {
    message("no accidents to plot")
    return(invisible(NULL))
  }
  is.na(data.sub$LONGITUD) <- data.sub$LONGITUD > 900
  is.na(data.sub$LATITUDE) <- data.sub$LATITUDE > 90
  with(data.sub, {
    maps::map("state", ylim = range(LATITUDE, na.rm = TRUE),
              xlim = range(LONGITUD, na.rm = TRUE))
    graphics::points(LONGITUD, LATITUDE, pch = 46)
  })
}


```
1. Data reading

Use fars_read() function in order to read csv file:

```{r echo = TRUE,results='asis', warnings=FALSE, messages = FALSE}

dt1<- fars_read(file.path("..","data-raw","accident_2013.csv.bz2"))
knitr::kable(head(as.data.frame(dt1)[,1:10],3))
```

2. Summarize data

Use fars_summarize_years() function in order to summarize data for a given years:


```{r  echo=TRUE, warnings=FALSE, messages = FALSE}


knitr::kable(fars_summarize_years(c(2014,2015)))
```

3. Map accidents

Use fars_map_state() function in order to map accidents for a given state:

``` {r echo = TRUE, warnings=FALSE, messages = FALSE}
fars_map_state(1,2015)
```

